---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(ggplot2)
```


```{r}

setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/")

outdir_files = "/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/files/"

outdir_figures = "/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/figures/"

sel_supertable_file = "files/selection_analyses_summary_table.txt"
sel_supertable = read.table(sel_supertable_file, sep = "\t", header = TRUE)

```
First we calculate the number of mutations in each bin:
1) Aggregated analysis w/ D-Loop
2) Aggregated analysis w/o D-Loop 
3) Analysis split per bin 

```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/")

per_bin_mut_counts_file = "files/total_mut_counts_all_bins_wo_dloop.txt"
per_bin_mut_counts = read.table(per_bin_mut_counts_file, header = TRUE, sep = "\t")

total_mut_count_wo_dloop_file = "../hnhs_wo_dloop/files/total_mut_count_wo_dloop.txt"
mut_counts_wo_dloop = read.table(total_mut_count_wo_dloop_file, header = TRUE, sep = "\t")

total_mut_count_w_dloop_file = "../files/mut_count_wo_hfps.txt"
mut_count_w_dloop = read.table(total_mut_count_w_dloop_file, header = TRUE, sep = "\t")

```
Preparing each dataframe to be merged -- we count the number of mutations in each bin: 

```{r}
per_bin_merge = per_bin_mut_counts %>%
  select(BIN, TOTAL_MUT_COUNT) %>%
  group_by(BIN) %>%
  summarise(COUNT = sum(TOTAL_MUT_COUNT))
```

```{r}
wo_dloop_merge = mut_counts_wo_dloop %>%
  mutate(BIN = "AGG_WO_DLOOP") %>%
  select(BIN, HET_COUNT) %>% 
  group_by(BIN) %>%
  summarise(COUNT = sum(HET_COUNT))

```

```{r}
w_dloop_merge = mut_count_w_dloop %>%
  mutate(BIN = "AGG_W_DLOOP") %>%
  select(BIN, HET_COUNT) %>% 
  group_by(BIN) %>%
  summarise(COUNT = sum(HET_COUNT))
```

Merging all of our mutation count information:
```{r}
mut_count_plotting_df = rbind(per_bin_merge, wo_dloop_merge, w_dloop_merge) %>%
  mutate(FACET = ifelse(grepl("AGG", BIN), "All", "Binned (- D-Loop)"))

mut_count_plotting_df$BIN = factor(mut_count_plotting_df$BIN, levels = c("AGG_W_DLOOP", "AGG_WO_DLOOP","upper-5e-5", "upper-1e-4", "upper-1e-3", "upper-1e-2", "upper-1"))
```

```{r}

rm(per_bin_merge, wo_dloop_merge, w_dloop_merge)

```

Plotting our mutation count distribution: 
https://stackoverflow.com/questions/52341385/how-to-automatically-adjust-the-width-of-each-facet-for-facet-wrap
```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/")

mut_count_dist = ggplot(mut_count_plotting_df, aes(x = BIN, y = COUNT)) + 
  geom_bar(stat = "identity") + 
  facet_grid(.~FACET, scales="free", space = "free") + 
  theme_bw(base_size = 14) + 
  scale_x_discrete(labels = c("AGG_W_DLOOP" = "+ D-Loop", "AGG_WO_DLOOP" = "- D-Loop","upper-5e-5" = "(0, 5e-5]", "upper-1e-4" = "(5e-5, 1e-4]", "upper-1e-3" = "(1e-4, 1e-3]", "upper-1e-2" = "(1e-3, 1e-2]", "upper-1" = "(1e-2, 1]")) + 
  xlab("Frequency bin") + 
  ylab("Mutation Counts") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 9.5),
        axis.title.x = element_text(size = 9.5),
        strip.background = element_blank(),
        strip.text.x = element_text(size = 10.25))

pdf(paste(outdir_figures,"/dist_muts_per_bin.pdf", sep=""),width=3.25,height=2.5)
print(mut_count_dist)
dev.off()


```
We calculate the proportion of significant hits per bin out of all the ratios we were able to calculate: 

```{r}

all_bins_poss_ratios = sel_supertable %>%
  select(BIN, POSSIBLE_RATIOS) %>%
  unique()

```

```{r}

sel_sig_counts = sel_supertable %>%
  #filter for our sig hits
  filter(P_ADJ < 0.01) %>%
  select(BIN, POSSIBLE_RATIOS) %>%
  group_by(BIN, POSSIBLE_RATIOS) %>%
  #within each bin how many sig hits do we have
  summarise(SIG_COUNT = n())

```

```{r}

sig_hit_signals_plotting = all_bins_poss_ratios %>%
  #to have all bins represented in the figure 
  left_join(sel_sig_counts, by = c("BIN", "POSSIBLE_RATIOS")) %>%
  ungroup() %>%
  mutate(SIG_COUNT =  ifelse(is.na(SIG_COUNT), 0, SIG_COUNT)) %>%
  mutate(PROP = SIG_COUNT/POSSIBLE_RATIOS) %>%
  mutate(FACET = ifelse(grepl("AGG", BIN), "All", "Binned (- D-Loop)"))

sig_hit_signals_plotting$BIN = factor(sig_hit_signals_plotting$BIN, levels = c("AGG_W_DLOOP", "AGG_WO_DLOOP","upper-5e-5", "upper-1e-4", "upper-1e-3", "upper-1e-2", "upper-1"))

```


```{r}

prop_sig_hits_plot = ggplot(sig_hit_signals_plotting, aes(x = BIN, y = PROP)) + 
  geom_bar(stat = "identity") + 
  facet_grid(.~FACET, scales="free", space = "free") + 
  theme_bw() + 
  scale_x_discrete(labels = c("AGG_W_DLOOP" = "+ D-Loop", "AGG_WO_DLOOP" = "- D-Loop", "upper-5e-5" = "(0, 5e-5]", "upper-1e-4" = "(5e-5, 1e-4]", "upper-1e-3" = "(1e-4, 1e-3]", "upper-1e-2" = "(1e-3, 1e-2]", "upper-1" = "(1e-2, 1]")) + 
  ylim(0, 0.3)+ 
  xlab("Frequency bin") + 
  ylab("Proportion \n[# of sig genes/# of ratios calculated]") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 9.5),
        axis.title.x = element_text(size = 9.5),
        strip.background = element_blank(),
        strip.text.x = element_text(size = 10.25))

pdf(paste(outdir_figures,"/prop_sig_hits_plot.pdf", sep=""),width=3,height=3)
print(prop_sig_hits_plot)
dev.off()

```

Type of selection within these bins

```{r}

type_sel_counts = sel_supertable %>%
  #filter our significant ratios 
  filter(P_ADJ < 0.01) %>%
  select(BIN, OBS_HNHS, P_ADJ) %>%
  mutate(SEL = ifelse(OBS_HNHS < 1, "N", "P")) %>%
  select(BIN, SEL) %>%
  group_by(BIN, SEL) %>%
  #counting the number of positive and negative selection in each bin
  summarise(COUNT = n())

```

```{r}
#making sure that each bin and selection combo is represented in the final figure
BIN = rep(c(all_bins_poss_ratios$BIN, "upper-1e-2", "upper-1"), each = 2)
SEL = rep(c("N", "P"), 7)

dummy_df = cbind(data.frame(BIN), data.frame(SEL))

```

Merging our dummy df and our selection df for plotting 

```{r}
plotting_sel_count = dummy_df %>%
  left_join(type_sel_counts, by = c("BIN", "SEL")) %>%
  mutate(COUNT = ifelse(is.na(COUNT), 0, COUNT)) %>%
  mutate(FACET = ifelse(grepl("AGG", BIN), "All", "Binned (- D-Loop)"))

plotting_sel_count$BIN = factor(plotting_sel_count$BIN, levels = c("AGG_W_DLOOP", "AGG_WO_DLOOP","upper-5e-5", "upper-1e-4", "upper-1e-3", "upper-1e-2", "upper-1"))
```

```{r}
type_sel = ggplot(plotting_sel_count, aes(x = BIN, y = COUNT, fill = SEL)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  facet_grid(.~FACET, scales="free", space = "free") + 
  theme_bw() + 
  scale_fill_manual(values = c("#a45851", "#5d74a5"), name = "Selection", labels = c("Negative", "Positive")) + 
  scale_x_discrete(labels = c("AGG_W_DLOOP" = "+ D-Loop", "AGG_WO_DLOOP" = "- D-Loop", "upper-5e-5" = "(0, 5e-5]", "upper-1e-4" = "(5e-5, 1e-4]", "upper-1e-3" = "(1e-4, 1e-3]", "upper-1e-2" = "(1e-3, 1e-2]", "upper-1" = "(1e-2, 1]")) + 
  xlab("Frequency bin") + 
  ylab("Count") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.y = element_text(size = 8.5),
        axis.title.x = element_text(size = 8.5),
        strip.background = element_blank(),
        strip.text.x = element_text(size = 9),
        legend.position = "none")

pdf(paste(outdir_figures,"/type_sel_plot.pdf", sep=""),width=3,height=2)
print(type_sel)
dev.off()

pdf(paste(outdir_figures,"/leg_type_sel_plot.pdf", sep=""),width=3.75,height=2.5)
print(type_sel + theme(legend.position = "bottom"))
dev.off()
```

```{r}

sel_supertable %>%
  filter(P_ADJ < 0.01) %>%
  filter(OBS_HNHS < 1)
```
