---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(ggplot2)
```

```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/strand_bias/")

outdir_figures = "figures/"
outdir_files = "files/"

#files that we use for our t-test comparisons -- includes the mutation frequency per sample
serrano_mutfreqs_file = "files/serrano_de_novo_mut_freqs.txt"
sc_mutfreqs_file = "files/sc_de_novo_mut_freqs.txt"


serrano_mutfreqs = read.table(serrano_mutfreqs_file, sep = '\t', header = TRUE)
sc_mutfreqs = read.table(sc_mutfreqs_file, sep = '\t', header = TRUE)

```
Creating our figure: 

```{r}
sc_plotting_df = sc_mutfreqs %>%
  select(MouseID, AGE_BIN, Tissue, MUT_TYPE, MUT_FREQ) %>%
  rename(SAMPLE = MouseID) %>%
  ungroup() %>%
  #we only compare the tissues that are also included in our study
  filter(Tissue == "L" | Tissue == "He" | Tissue == "C") %>%
  mutate(TISSUE_LABEL = recode(Tissue, C = "WB/Cere", He = "Heart", L = "Liver")) %>%
  mutate(AGE_LABEL = ifelse(AGE_BIN == "OLD", "Old", "Young")) %>%
  select(SAMPLE, TISSUE_LABEL, AGE_LABEL, MUT_TYPE, MUT_FREQ) %>%
  group_by(TISSUE_LABEL, AGE_LABEL, MUT_TYPE) %>%
  mutate(AVG = mean(MUT_FREQ), SAMP_COUNT = n(), SE = sd(MUT_FREQ)/sqrt(SAMP_COUNT)) %>%
  mutate(STUDY = "Sanchez-Contreras") %>%
  select(SAMPLE, TISSUE_LABEL, AGE_LABEL, MUT_TYPE, MUT_FREQ, AVG, SAMP_COUNT, SE, STUDY)
  
```

```{r}

serrano_plotting_df = serrano_mutfreqs %>%
  #we only compare across B6 ancestry, which is what SC have in their study
  filter(STRAIN == "B6") %>%
  #recoding the brain and age label to match that of SC
  mutate(TISSUE_LABEL = recode(TISSUE, Brain = "WB/Cere")) %>%
  mutate(AGE_LABEL = ifelse(AGE_BIN == "OLD", "Old", "Young")) %>%
  select(SAMPLE, TISSUE_LABEL, AGE_LABEL, MUT_TYPE, MUT_FREQ) %>%
  group_by(TISSUE_LABEL, AGE_LABEL, MUT_TYPE) %>%
  mutate(AVG = mean(MUT_FREQ), SAMP_COUNT = n(), SE = sd(MUT_FREQ)/sqrt(SAMP_COUNT)) %>%
  #adding the study name 
  mutate(STUDY = "Serrano") 

```


Merging our datasets to plot study comparisons

```{r}
comp_df = serrano_plotting_df %>% 
  rbind(sc_plotting_df) %>%
  mutate(DUMMY_X = case_when(
    MUT_TYPE == "G>A" ~ 1, MUT_TYPE == "C>T" ~ 2,
    MUT_TYPE == "G>C" ~ 3, MUT_TYPE == "C>G" ~ 4,
    MUT_TYPE == "G>T" ~ 5, MUT_TYPE == "C>A" ~ 6,
    MUT_TYPE == "T>A" ~ 7, MUT_TYPE == "A>T" ~ 8,
    MUT_TYPE == "T>C" ~ 9, MUT_TYPE == "A>G" ~ 10,
    MUT_TYPE == "T>G" ~ 11, MUT_TYPE == "A>C" ~ 12
  )) %>%
  mutate(DUMMY_X_POINT = ifelse(STUDY == "Serrano", DUMMY_X + 0.23, DUMMY_X - 0.23)) %>%
  mutate(DUMMY_X_POINT = ifelse((TISSUE_LABEL == "Liver" & AGE_LABEL == "Young"), DUMMY_X, DUMMY_X_POINT))

comp_df$MUT_TYPE = factor(comp_df$MUT_TYPE, level = c("G>A","C>T", "G>C", "C>G", "G>T", "C>A", "T>A", "A>T", "T>C", "A>G", "T>G", "A>C"))
comp_df$AGE_LABEL = factor(comp_df$AGE_LABEL, level = c("Young", "Old"))
comp_df$TISSUE_LABEL = factor(comp_df$TISSUE_LABEL, level = c("Heart", "Liver", "WB/Cere"))
```

Plotting our frequencies against each other

```{r}

setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/strand_bias/")

study_freq_comp = ggplot(comp_df, aes(x = DUMMY_X, y = AVG, fill = STUDY)) 

study_freq_comp = study_freq_comp + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin=AVG-SE, ymax=AVG+SE), width=.2,
                 position=position_dodge(.9)) +
  geom_point(aes(x = DUMMY_X_POINT, y = MUT_FREQ), color = "black", size = 0.4, alpha = 0.7) + 
  facet_grid(TISSUE_LABEL ~ AGE_LABEL, scales = "free_y") +
  theme_bw(base_size = 12) + 
  xlab("Mutation Type") + 
  ylab("Mutation frequency") + 
  scale_x_continuous(breaks = 1:12, labels = c("G>A","C>T", "G>C", "C>G", "G>T", "C>A", "T>A", "A>T", "T>C", "A>G", "T>G", "A>C")) + 
  scale_fill_manual(values = c("#A9A9A9","#9F2B68")) +
  theme(axis.text.x = element_text(angle = 90), 
        strip.background = element_blank(),
        legend.position = "bottom")
  

pdf(paste(outdir_figures,"/study_comp_freqs.pdf", sep=""),width=6,height=4)
print(study_freq_comp)
dev.off()

```

Now statistically testing if there's a difference between our mutation frequencies: 

Heart, Liver, and Cere for each mutation type

```{r}

serrano_t_df = serrano_mutfreqs %>%
  filter(STRAIN == "B6") %>%
  mutate(LABEL = paste(TISSUE, AGE_BIN, sep = "_")) %>% 
  select(LABEL, MUT_TYPE, MUT_FREQ)

```

```{r}
sc_t_df = sc_mutfreqs %>%
  filter(Tissue == "He" | Tissue == "L" | Tissue == "C") %>%
  mutate(TISSUE = recode(Tissue, "He" = "Heart", "L" = "Liver", "C" = "Brain")) %>% 
  mutate(LABEL = paste(TISSUE, AGE_BIN, sep = "_")) %>% 
  #filter our young liver, since we do not have this condition in our study
  filter(LABEL != "Liver_YOUNG") %>% 
  select(LABEL, MUT_TYPE, MUT_FREQ)
```

The lists that we will be iterating through: 

```{r}
COND_LABEL = unique(serrano_t_df$LABEL)
MUT_TYPE = unique(serrano_t_df$MUT_TYPE)
```

Now to loop through our dataframes and run a t-test on all mutation type combinations 

```{r}
#initialize t-test results df 
t_output = data.frame(CONDITION = character(),
                      MUT_TYPE = character(),
                      P_VAL = numeric(),
                      STAT = numeric()
)

#we want to loop through the label (tissue x age combination)
for(label in COND_LABEL){
  
#filter df based on the label 
  serrano_cond_df = serrano_t_df %>%
    filter(LABEL == label)
  
    sc_cond_df = sc_t_df %>%
    filter(LABEL == label)
  
#we want to loop through the mutation types
    
    for(type in MUT_TYPE){
      
      #create vectors of the frequencies for the corresponding mutation type
      serrano_entries = serrano_cond_df[serrano_cond_df$MUT_TYPE == type,]$MUT_FREQ
      sc_entries = sc_cond_df[sc_cond_df$MUT_TYPE == type,]$MUT_FREQ
      
      #now run a t-test
      t_test_results = t.test(serrano_entries, sc_entries)
      
      pval = t_test_results$p.value
      stat = t_test_results$statistic
    
    #save the condition label, the mutation pair label, p-value, and statistic
    output_results = c(label, type, pval, stat)
    
    #save results onto new row of the output table 
    t_output[nrow(t_output) + 1,] = output_results
}}

```

Correcting for multiple hypothesis testing

```{r}
adj_t_output = t_output %>%
  mutate(P_ADJ = p.adjust(P_VAL, method="BH")) %>%
  arrange(P_ADJ)
```

Exporting the t-test results
```{r}

outdir_files ="/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/strand_bias/files/"  

#exporting our strand bias fishers exact test 
write.table(adj_t_output, file = paste(outdir_files,"/serrano_v_sc_t_test.txt", sep = ""), sep = "\t", quote = F, row.names = F)

```

Testing our results -- all clear! 
```{r}

serrano_test_cond = serrano_t_df %>%
    filter(LABEL == "Heart_YOUNG")

sc_test_cond = sc_t_df %>%
    filter(LABEL == "Heart_YOUNG")


serrano_test = serrano_test_cond[serrano_test_cond$MUT_TYPE == "G>T",]$MUT_FREQ
sc_test = sc_test_cond[sc_test_cond$MUT_TYPE == "G>T",]$MUT_FREQ

t.test(serrano_test, sc_test)

```

Lastly, we want to compare the asymmetry between our studies

```{r}

serrano_assym = serrano_mutfreqs %>%
  filter(STRAIN == "B6") %>%
  mutate(COL_LABELS = recode(MUT_TYPE, "G>A" = "G_A", "C>T" = "C_T")) %>%
  filter(MUT_TYPE == "G>A" | MUT_TYPE == "C>T") %>%
  select(SAMPLE, TISSUE, AGE_BIN, COL_LABELS, MUT_FREQ) %>% 
  pivot_wider(names_from = COL_LABELS, values_from = MUT_FREQ) %>%
  mutate(RATIO = G_A/C_T) %>%
  ungroup() %>%
  select(TISSUE, AGE_BIN, RATIO) %>%
  group_by(TISSUE, AGE_BIN) %>%
  summarise(AVG_RATIO = mean(RATIO), COUNT = n(), SE = sd(RATIO)/sqrt(COUNT)) %>%
  mutate(Tissue_Label = recode(TISSUE, "Brain" = "WB/C", "Heart" = "H", "Liver" = "L")) %>%
  mutate(Study = "Serrano") %>%
  ungroup() %>%
  select(Study, AGE_BIN, Tissue_Label, AVG_RATIO, SE)

```

```{r}

sc_assym = sc_mutfreqs %>%
  select(MouseID, AGE_BIN, Tissue, MUT_TYPE, MUT_FREQ) %>%
  filter(Tissue == "He" | Tissue == "L" | Tissue == "C") %>% 
  filter(MUT_TYPE == "G>A" | MUT_TYPE == "C>T") %>%
  mutate(COL_LABELS = recode(MUT_TYPE, "G>A" = "G_A", "C>T" = "C_T")) %>%
  select(MouseID, AGE_BIN, Tissue, COL_LABELS, MUT_FREQ) %>% 
  pivot_wider(names_from = COL_LABELS, values_from = MUT_FREQ) %>%
  mutate(RATIO = G_A/C_T) %>%
  ungroup() %>%
  select(Tissue, AGE_BIN, RATIO) %>%
  group_by(Tissue, AGE_BIN) %>%
  summarise(AVG_RATIO = mean(RATIO), COUNT = n(), SE = sd(RATIO)/sqrt(COUNT)) %>%
  mutate(Tissue_Label = recode(Tissue, "C" = "WB/C", "He" = "H", "L"= "L")) %>%
  mutate(Study = "Sanchez-Contreras") %>% 
  ungroup() %>%
  select(Study, AGE_BIN, Tissue_Label, AVG_RATIO, SE)
```

```{r}

assym_comp_df = rbind(serrano_assym, sc_assym) %>%
  mutate(AGE_LABEL = ifelse(AGE_BIN == "OLD", "Old", "Young"))

assym_comp_df$AGE_LABEL = factor(assym_comp_df$AGE_LABEL, level = c("Young", "Old"))

```

```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/strand_bias/")

assym_comp_fig = ggplot(assym_comp_df, aes(x = Tissue_Label, y = AVG_RATIO, group = Study, color = Study))

assym_comp_fig = assym_comp_fig + 
  geom_point() + 
  facet_wrap(AGE_LABEL~., nrow = 1) + 
  geom_pointrange(aes(ymin=AVG_RATIO-SE, ymax=AVG_RATIO+SE)) + 
  theme_bw(base_size = 16) + 
  scale_color_manual(values = c("#A9A9A9","#9F2B68")) +
  ylab("Average ratio of mutation frequencies \n(G>A/C>T)") + 
  xlab("Tissue") + 
  theme(strip.background = element_blank(),
        strip.text.x = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        legend.position = "bottom")
  
pdf(paste(outdir_figures,"study_comp_strand_assym.pdf", sep=""),width=5,height=6)
print(assym_comp_fig)
dev.off()
```
