---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(ggplot2)
```


```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/")

outdir_files =  "/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/files/"
outdir_figures =  "/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/figures/"

supertable_file ="/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/input_files/supertable.txt"
supertable = read.table(supertable_file, header=TRUE, stringsAsFactors = FALSE)
```

Creating the mutation count without D-Loop file 
```{r}
total_mut_count_wo_dloop = supertable %>%
  #filter for SNVs 
  filter(VARIANT_TYPE == "SNV") %>%
  #we are removing the D-Loop which we know is a hotspot in the mt-genome 
  filter(GENE != "D-Loop") %>%
  #filter out mutations that are segregating at high frequency
  filter(CONDITION_MUT_FREQ_AT_POS < 0.001) %>%
  #calculating the mutation count within a condition
  select(STRAIN, TISSUE, AGE_BIN, ALT_ALLELE_DEPTH) %>%
  group_by(STRAIN, TISSUE, AGE_BIN) %>%
  summarise(HET_COUNT = sum(ALT_ALLELE_DEPTH)) %>%
  #this is needed so that the FVB strain name matches across files -- ugh
  mutate(STRAIN = recode(STRAIN, "FVB" = "F"))

```

```{r}
outdir_files ="/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/files/"  

#exporting our strand bias fishers exact test 
write.table(total_mut_count_wo_dloop, file = paste(outdir_files,"/total_mut_count_wo_dloop.txt", sep = ""), sep = "\t", quote = F, row.names = F)
```

We need to calculate the proportion of each mutation type -- this will give us the probability of drawing a mutation of each type 
```{r}
mut_type_counts = supertable %>%
  #filter for SNVs 
  filter(VARIANT_TYPE == "SNV") %>%
  #we are removing the D-Loop which we know is a hotspot in the mt-genome 
  filter(GENE != "D-Loop") %>% 
  #filter out mutations that are segregating at high frequency
  filter(CONDITION_MUT_FREQ_AT_POS < 0.001) %>%
  mutate(MUTATION_TYPE = paste(REF, ALT, sep = ">")) %>%
  select(STRAIN, TISSUE, AGE_BIN, MUTATION_TYPE, ALT_ALLELE_DEPTH) %>%
  group_by(STRAIN, TISSUE, AGE_BIN, MUTATION_TYPE) %>%
  summarise(MUT_TYPE_COUNT = sum(ALT_ALLELE_DEPTH)) %>%
  #this is needed so that the FVB strain name matches across files -- ugh
  mutate(STRAIN = recode(STRAIN, "FVB" = "F"))
```

Creating a dummy df to make sure we account for mut types that do not appear in a condition
```{r}

condition = unique((mut_type_counts %>%
  select(STRAIN, TISSUE, AGE_BIN) %>%
  mutate(COND = paste(STRAIN, TISSUE, AGE_BIN, sep = "_")))$COND)

dummy_cond = rep(condition, each = 12)

dummy_mut_type = rep(unique(mut_type_counts$MUTATION_TYPE), 29)

dummy_df = cbind(data.frame(dummy_cond), data.frame(dummy_mut_type)) %>%
  separate(dummy_cond, into = c("STRAIN", "TISSUE", "AGE_BIN")) %>%
  rename(MUTATION_TYPE = dummy_mut_type)
```

Merging the dummy df with the mut counts data frame to account for all mutation types across conditions
```{r}

#there are 2 conditions that are missing a mutation type 
mut_type_count_all_muts_wo_dloop = dummy_df %>%
  left_join(mut_type_counts, by = c("STRAIN", "TISSUE", "AGE_BIN", "MUTATION_TYPE")) %>%
  mutate(MUT_TYPE_COUNT = ifelse(is.na(MUT_TYPE_COUNT), 0, MUT_TYPE_COUNT))

```

Now we can calculate the proportion of mutations in the non-dloop regions

```{r}
mut_type_props_nondloop = mut_type_count_all_muts_wo_dloop %>%
  left_join(total_mut_count_wo_dloop, by = c("STRAIN", "TISSUE", "AGE_BIN")) %>%
  mutate(HET_PROP = MUT_TYPE_COUNT/HET_COUNT) %>%
  select(STRAIN, TISSUE, AGE_BIN, MUTATION_TYPE, MUT_TYPE_COUNT, HET_PROP) %>%
  mutate(REGION = "NON_DLOOP")
```

```{r}
#exporting the file 
outdir_files ="/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/files/"  

#exporting our strand bias fishers exact test 
write.table(mut_type_props_nondloop, file = paste(outdir_files,"/mut_type_props_nondloop.txt", sep = ""), sep = "\t", quote = F, row.names = F)

```

To calculate the proportions of mutation types in the D-Loop we repeat the analyses above for just the mutation in the D-Loop region

```{r}

total_mut_count_in_dloop = supertable %>%
  #filter for SNVs 
  filter(VARIANT_TYPE == "SNV") %>%
  #we are removing the D-Loop which we know is a hotspot in the mt-genome 
  filter(GENE == "D-Loop") %>% 
  #filter out mutations that are segregating at high frequency
  filter(CONDITION_MUT_FREQ_AT_POS < 0.001) %>%
  #calculating the mutation count within a condition
  select(STRAIN, TISSUE, AGE_BIN, ALT_ALLELE_DEPTH) %>%
  group_by(STRAIN, TISSUE, AGE_BIN) %>%
  summarise(HET_COUNT = sum(ALT_ALLELE_DEPTH)) %>%
  #this is needed so that the FVB strain name matches across files -- ugh
  mutate(STRAIN = recode(STRAIN, "FVB" = "F"))

```

Calculating the mutation type count in the D-Loop

```{r}
mut_types_counts_in_dloop = supertable %>%
  #filter for SNVs 
  filter(VARIANT_TYPE == "SNV") %>%
  #we are removing the D-Loop which we know is a hotspot in the mt-genome 
  filter(GENE == "D-Loop") %>% 
  #filter out mutations that are segregating at high frequency
  filter(CONDITION_MUT_FREQ_AT_POS < 0.001) %>%
  mutate(MUTATION_TYPE = paste(REF, ALT, sep = ">")) %>%
  select(STRAIN, TISSUE, AGE_BIN, MUTATION_TYPE, ALT_ALLELE_DEPTH) %>%
  group_by(STRAIN, TISSUE, AGE_BIN, MUTATION_TYPE) %>%
  summarise(MUT_TYPE_COUNT = sum(ALT_ALLELE_DEPTH)) %>%
  #this is needed so that the FVB strain name matches across files -- ugh
  mutate(STRAIN = recode(STRAIN, "FVB" = "F"))

```
We want to make sure we account for mutation types that did not occur in the D-Loop 
```{r}

mut_type_props_dloop = dummy_df %>%
  left_join(mut_types_counts_in_dloop, by = c("STRAIN", "TISSUE", "AGE_BIN", "MUTATION_TYPE")) %>%
  mutate(MUT_TYPE_COUNT = ifelse(is.na(MUT_TYPE_COUNT), 0, MUT_TYPE_COUNT)) %>%
  left_join(total_mut_count_in_dloop, by = c("STRAIN", "TISSUE", "AGE_BIN")) %>%
  mutate(HET_PROP = MUT_TYPE_COUNT/HET_COUNT) %>%
  select(STRAIN, TISSUE, AGE_BIN, MUTATION_TYPE, MUT_TYPE_COUNT, HET_PROP) %>%
  mutate(REGION = "DLOOP") 
  

```


Now to compare the D-Loop v non-D-Loop mutational landscapes 

```{r}
comparing_props_df = mut_type_props_dloop %>%
  rbind(mut_type_props_nondloop) %>%
  select(STRAIN, TISSUE, AGE_BIN, MUTATION_TYPE, HET_PROP, REGION) 
  
```

```{r}
summary_prop_comp_df = comparing_props_df %>%
  select(TISSUE, AGE_BIN, MUTATION_TYPE, HET_PROP, REGION) %>%
  group_by(TISSUE, AGE_BIN, MUTATION_TYPE, REGION) %>%
  summarise(AVG = mean(HET_PROP), N = n(), SE = sd(HET_PROP)/ sqrt(N))

#rearranging our mutation based on frequency 
summary_prop_comp_df$MUTATION_TYPE = factor(summary_prop_comp_df$MUTATION_TYPE, levels = c("G>A", "C>T", "G>T", "C>A", "T>C", "A>G", "G>C", "C>G", "T>A", "A>T", "T>G", "A>C"))
summary_prop_comp_df$AGE_BIN = factor(summary_prop_comp_df$AGE_BIN, levels = c("YOUNG", "OLD"))
```

Plotting the comparison of mutation proportions between D-Loop and non-D-Loop

```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/")
age.labels = c("Young", "Old")
names(age.labels) = c("YOUNG", "OLD")

comp_props_fig = ggplot(summary_prop_comp_df, aes(x = MUTATION_TYPE, y = AVG, color = REGION))

comp_props_fig = comp_props_fig + 
  geom_point(size = 0.5) + 
  geom_errorbar(aes(ymin=AVG-SE, ymax=AVG+SE), width = 0.2) + 
  facet_grid(TISSUE ~ AGE_BIN, labeller = labeller(AGE_BIN = age.labels)) + 
  theme_bw(base_size = 16) + 
  xlab("Mutation Type") + 
  ylab("Average Proportion") + 
  scale_color_manual(name = "Region", labels = c("D-Loop", "Non D-Loop"), values = c("#A9A9A9","#9F2B68")) +
  theme(strip.background = element_blank(), 
        axis.text.x = element_text(angle = 90, size = 11),
        axis.text.y = element_text(size = 11),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        legend.position = "bottom")

pdf(paste(outdir_figures,"/dloop_v_nondloop_mut_props.pdf", sep=""),width=5,height=4)
print(comp_props_fig)
dev.off()


```
