---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(ggplot2)
```

```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/")

outdir_figures = "/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/strand_bias/figures/"
outdir_files = "/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/strand_bias/files/"

sc_supertable_file = "~/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/sc2023_mouse_aging/files/sc2023_mouse_aging_supertable.txt"

sc_summary_file = "~/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/sc2023_mouse_aging/files/Mouse_aging_mtDNA_summary.csv"

#note specifying the delimeter eliminates the rows with ## or # in the entry -- the headers that were not filtered out because these 3 files had 2 more header lines than the other files 
sc_supertable = read.table(sc_supertable_file, sep = '\t', header = TRUE)

sc_summary = read.csv(sc_summary_file, header = TRUE )
```

```{r}

sc_supertable %>%
  filter(Treatment == "NT")
  

```

Let's find the NT samples and the DP depth for each nucleotide per sample 

1. Checking the ratio of Cs to Gs -- ranges from ~1.93 - 2 
```{r}
sc_summary %>%
  filter(Treatment == "NT") %>% 
  select(MouseID, Age, Treatment, Tissue, C.s_Sequenced, G.s_Sequenced) %>%
  mutate(RATIO = C.s_Sequenced/G.s_Sequenced)
  
```
2. Creating our dataframe for merging 
```{r}

sc_duplex_ref = sc_summary %>%
  filter(Treatment == "NT") %>% 
  select(MouseID, Age, Treatment, Tissue, A.s_Sequenced, T.s_Sequenced, C.s_Sequenced, G.s_Sequenced) %>%
  pivot_longer(cols = A.s_Sequenced:G.s_Sequenced, names_to = "NUC_TYPE", values_to = "DP_BP") %>% 
  mutate(REF = recode(NUC_TYPE, A.s_Sequenced = "A",
                      T.s_Sequenced = "T",
                      C.s_Sequenced = "C",
                      G.s_Sequenced = "G")) %>%
  select(MouseID, Age, Treatment, Tissue, REF, DP_BP) %>%
  mutate(AGE_BIN = ifelse(Age == 26, "OLD", "YOUNG")) %>%
  select(MouseID, Tissue, AGE_BIN, REF, DP_BP)
  

```

Now, we count the de novo mutations

```{r}
sc_de_novo_muts = sc_supertable %>%
  filter(Treatment == "NT") %>% 
  #this is where we start to filter for de novo mutations
  #we remove ~200 mutations 
  filter(ALT_ALLELE_DEPTH < 100) %>%
  #calculate the mutation frequency at each position for each sample
  mutate(MUT_FREQ = ALT_ALLELE_DEPTH/READ_DEPTH_AT_POS) %>%
  #filter any mutation with a frequency > 1% 
  filter(MUT_FREQ < 0.01) %>%
  #filter out any position that did not have a mutation
  filter(MUT_FREQ > 0) %>%
  mutate(AGE_BIN = ifelse(Age == 26, "OLD", "YOUNG")) %>%
  select(MouseID, AGE_BIN, Tissue, POS, REF, ALT) %>%
  #scoring each mutation type once 
  unique() %>%
  #filtering for SNVs
  filter(nchar(REF) == 1) %>%
  filter(nchar(ALT) == 1) %>%
  mutate(MUT_TYPE = paste(REF, ALT, sep = ">")) %>%
  select(MouseID, AGE_BIN, Tissue, MUT_TYPE) %>%
  group_by(MouseID, AGE_BIN, Tissue, MUT_TYPE) %>%
  summarise(COUNT = n())
```

```{r}
sc_de_novo_muts %>%
  ungroup() %>%
  select(MouseID, Tissue) %>%
  unique() %>%
  mutate(LABEL = paste(MouseID, Tissue, sep = "_"))

```

We need to create a dummy df for mutation types that had a 0 mutation frequency 

```{r}

sample_names = rep(((sc_de_novo_muts %>%
  ungroup() %>%
  select(MouseID, Tissue, AGE_BIN) %>%
  unique() %>%
  mutate(LABEL = paste(MouseID, Tissue, AGE_BIN, sep = "_")))$LABEL), each = 12)

mut_types = rep(((sc_de_novo_muts %>%
  ungroup() %>%
  select(MUT_TYPE) %>%
  unique())$MUT_TYPE), 88)

dummy_df = cbind(data.frame(sample_names), data.frame(mut_types)) %>%
  separate(sample_names, sep = "_", into = c("MouseID", "Tissue", "AGE_BIN"))

colnames(dummy_df) = c("MouseID", "Tissue", "AGE_BIN", "MUT_TYPE")

```

Now to merge our dummy df and our de novo counts 

```{r}

sc_mut_freqs = sc_de_novo_muts %>%
  ungroup() %>%
  full_join(dummy_df, by = c("MouseID", "Tissue", "AGE_BIN", "MUT_TYPE")) %>%
  mutate(COUNT = ifelse(is.na(COUNT), 0, COUNT)) %>%
  separate(MUT_TYPE, sep = ">", into = c("REF", "ALT"), remove = FALSE) %>%
  select(MouseID, AGE_BIN, Tissue, MUT_TYPE, REF, COUNT) %>%
  left_join(sc_duplex_ref, by = c("MouseID", "AGE_BIN", "Tissue", "REF")) %>%
  mutate(MUT_FREQ = COUNT/DP_BP)

```

Now to create our summary df 

```{r}

sc_summary_df = sc_mut_freqs %>%
  ungroup() %>%
  select(Tissue, AGE_BIN, MUT_TYPE, MUT_FREQ) %>%
  group_by(Tissue, AGE_BIN, MUT_TYPE) %>%
  summarise(AVG = mean(MUT_FREQ), SAMP_COUNT = n(), SE = sd(MUT_FREQ)/sqrt(SAMP_COUNT))

sc_summary_df$MUT_TYPE = factor(sc_summary_df$MUT_TYPE, level = c("G>A","C>T", "G>C", "C>G", "G>T", "C>A", "T>A", "A>T", "T>C", "A>G", "T>G", "A>C"))
sc_summary_df$AGE_BIN = factor(sc_summary_df$AGE_BIN, level = c("YOUNG", "OLD"))
```

Plotting our mut freqs!

```{r}

summary_sc_plot = ggplot(sc_summary_df, aes(x = MUT_TYPE, y = AVG, fill = Tissue)) 

summary_sc_plot = summary_sc_plot + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin=AVG-SE, ymax=AVG+SE), width=.2,
                 position=position_dodge(.9)) +
  facet_wrap(. ~ AGE_BIN, nrow = 2) + 
  ylab("Mutation frequency") + 
  xlab("Mutation Type") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90), 
        legend.position = "bottom")

pdf(paste(outdir_figures,"/sc_frequencies_strand_bias.pdf",sep=""),width=6,height=4 )
print(summary_sc_plot)
dev.off()
  
```

Exporting our mutation frequencies 
```{r}
outdir_files ="/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/strand_bias/files/"  

#exporting our strand bias fishers exact test 
write.table(sc_mut_freqs, file = paste(outdir_files,"/sc_de_novo_mut_freqs.txt", sep = ""), sep = "\t", quote = F, row.names = F)

#exporting our strand bias fishers exact test 
write.table(sc_summary_df, file = paste(outdir_files,"/sc_summary_de_novo_mut_freqs.txt", sep = ""), sep = "\t", quote = F, row.names = F)

```

Running a t-test analysis for strand bias

Now we want to statistically test if there's a difference between the complementary average mutation frequencies for each condition

```{r}
t_sc_df = sc_mut_freqs %>%
  mutate(COND_LABEL = paste(Tissue, AGE_BIN, sep = "_"))
```

The list of mutation pairs and cond_labels we will iterate through
```{r}
MUT_PAIRS = c("G>A_C>T","G>C_C>G", "G>T_C>A",
              "A>C_T>G", "A>G_T>C", "A>T_T>A")


COND_LABEL = (t_sc_df %>%
  select(COND_LABEL) %>%
  unique())$COND_LABEL

```

```{r}
#initialize our t-test results df 
t_output = data.frame(CONDITION = character(),
                      MUT_PAIR = character(),
                      P_VAL = numeric(),
                      T_STAT = numeric()
)

#loop through each condition in one loop via the LABEL field 
for(label in COND_LABEL){
  #subset our mutation freq df for just this label 
  print(label)
  
  cond_subset_df = t_sc_df %>%
    filter(COND_LABEL == label)
  
  #loop through this subset for our specific mutation pairs that we can find via the PAIRS field
  for(pair in MUT_PAIRS){
    #break down the pair label based on the delim 
    #assign the mutation and compliment to the variables 
    #print variables 
    print(pair) 
    split = str_split(pair, "_")

    #sigh totally breaking an abstraction here
    mut = split[[1]][1]
    comp_mut = split[[1]][2]
    
    print(mut)
    print(comp_mut)
    
    #subset a df that has these mutation types 
    t_test_df = cond_subset_df %>%
      filter(MUT_TYPE == mut | MUT_TYPE == comp_mut)
    
    #run the t-test on this subset df
    t_test_results = t.test(MUT_FREQ~MUT_TYPE,data=t_test_df)
    
    pval = t_test_results$p.value
    stat = t_test_results$statistic
    
    #save the condition label, the mutation pair label, p-value, and statistic
    output_results = c(label, pair, pval, stat)
    
    
    #save results onto new row of the output table 
    t_output[nrow(t_output) + 1,] = output_results
  }
}


```
Multiple hypothesis correcting 

```{r}
corrected_t_output = t_output %>%
  mutate(P_ADJ = p.adjust(P_VAL, method="BH"))

```

```{r}

corrected_t_output %>%
  filter(P_ADJ < 0.01) 

```

We are running the strand bias analysis with a Fisher's Exact test, since our power comes from the mutation counts in this statistical test and we test for dependence between mut type and strand

```{r}

cond_sc_mut_count = sc_mut_freqs %>%
  select(Tissue, AGE_BIN, MUT_TYPE, COUNT) %>%
  group_by(Tissue, AGE_BIN, MUT_TYPE) %>%
  summarise(COND_MUT_COUNT = sum(COUNT)) %>%
  mutate(COND_LABEL = paste(Tissue, AGE_BIN, sep = "_"))

```

We need the duplex depth for the Fisher's Exact 

```{r}
sc_cond_duplex_depth = sc_duplex_ref %>%
  select(Tissue, AGE_BIN, REF, DP_BP) %>%
  mutate(COND_LABEL = paste(Tissue, AGE_BIN, sep = "_")) %>%
  select(COND_LABEL, REF, DP_BP) %>%
  group_by(COND_LABEL, REF) %>%
  summarise(DP_SUM = sum(DP_BP))


```

Exporting these two files for the study comparison analysis

```{r}

outdir_files ="/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/strand_bias/files/"  

#exporting our strand bias fishers exact test 
write.table(cond_sc_mut_count, file = paste(outdir_files,"/sc_cond_mut_counts.txt", sep = ""), sep = "\t", quote = F, row.names = F)

write.table(sc_cond_duplex_depth, file = paste(outdir_files,"/sc_cond_ref_bases.txt", sep = ""), sep = "\t", quote = F, row.names = F)

```

Now creating a for loop for our fishers exact test

```{r}
#initializing our df 
fe_output = data.frame(COND_LABEL = character(),
                    MUT_PAIR = character(),
                    P_VAL = numeric(),
                    OR = numeric()) 

#looping through our mutation pairs 
for(pair in MUT_PAIRS){
    #break down the pair label based on the delim 
    #assign the mutation and compliment to the variables 
    #print variables 
    print(pair) 
    split = str_split(pair, "_")

    #sigh totally breaking an abstraction here
    mut = split[[1]][1]
    comp_mut = split[[1]][2]
  
    #split the ref based on the mut 
    mut_ref = str_split(mut, ">")[[1]][1]
    comp_ref = str_split(comp_mut, ">")[[1]][1]
    
    #filter data based on mut and comp mut
    mut_pair_subset_df = cond_sc_mut_count %>%
      filter(MUT_TYPE == mut | MUT_TYPE == comp_mut)
    
    #now loop through each cond_label 
    for(label in COND_LABEL){
      #filter the duplex_bp data based on the cond_label 
      duplex_bp_subset = sc_cond_duplex_depth %>%
        filter(COND_LABEL == label)
      
    #subset df for this label 
      
      contigency_df_inputs = mut_pair_subset_df %>%
        filter(COND_LABEL == label)
      
      #grab items for this condition fe contigency table 
      mut_counts = contigency_df_inputs[contigency_df_inputs$MUT_TYPE == mut,]$COND_MUT_COUNT
      comp_mut_counts = contigency_df_inputs[contigency_df_inputs$MUT_TYPE == comp_mut,]$COND_MUT_COUNT
      
      mut_ref_counts = duplex_bp_subset[duplex_bp_subset$REF == mut_ref,]$DP_SUM
      comp_ref_counts = duplex_bp_subset[duplex_bp_subset$REF == comp_ref,]$DP_SUM
      
      #create contigency table
      con_tbl = matrix(c(mut_counts, mut_ref_counts - mut_counts, comp_mut_counts, comp_ref_counts - comp_mut_counts), ncol = 2, byrow = F)
      
      #running fisher's exact on the df 
      fishers_results = fisher.test(con_tbl, alternative = "two.sided")

      pval = fishers_results$p.value[1]
      OR = fishers_results$estimate

      #appending to our df 
      results = c(label, pair, pval, OR)
  
      fe_output[nrow(fe_output) + 1,] = results
    
}}

```

```{r}

corrected_fe_output = fe_output %>%
  mutate(P_ADJ = p.adjust(P_VAL, method="BH"))
  

```

```{r}
corrected_fe_output %>%
  filter(P_ADJ < 0.01)

```
