---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(ggplot2)
```

```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/")

outdir_files =  "/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/files/"
outdir_figures = "/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/figures"

#loading in our simulations 
sims_file = "files/combined_simulations_counts_per_gene.txt"
sims = read.table(sims_file, header=TRUE, stringsAsFactors = FALSE)
```

```{r}
sims_wider = sims %>%
  mutate(STRAIN = ifelse(STRAIN == "F", "FVB", STRAIN)) %>%
  pivot_wider(names_from = ANNOTATION, values_from = COUNT)

```

```{r}
rm(sims)
```

Drop the simulations where we would not be able to calculate an hn/hs ratio due to 0 nonsyn or 0 syn counts
Relabel the FVB strain for merging and relabel the nonsyn and syn count columns accordingly 
```{r}
filtered_sims_wider = sims_wider %>%
  drop_na() %>%
  rename(SIM_NONSYN_COUNT = missense_variant, SIM_SYN_COUNT = synonymous_variant) 
```

```{r}
rm(sims_wider)
```

Import our total_num_sites file 

```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/")

total_num_sites_file = "files/total_annotated_num_sites_per_gene.txt"
total_num_sites = read.table(total_num_sites_file, header=TRUE, stringsAsFactors = FALSE)

```

Merging our denominator and numerator files

```{r}
sim_ratios = filtered_sims_wider %>%
  left_join(total_num_sites, by = c("STRAIN", "TISSUE", "AGE_BIN", "GENE")) %>%
  select(STRAIN, TISSUE, AGE_BIN, BIN, SIM_RUN, GENE, SIM_NONSYN_COUNT, SIM_SYN_COUNT, TOTAL_NONSYN_SITE_COUNT, TOTAL_SYN_SITE_COUNT) %>%
  mutate(hN = SIM_NONSYN_COUNT/TOTAL_NONSYN_SITE_COUNT, hS = SIM_SYN_COUNT/TOTAL_SYN_SITE_COUNT) %>%
  mutate(SIM_RATIO = hN/hS) %>%
  select(STRAIN, TISSUE, AGE_BIN, BIN, SIM_RUN, GENE, SIM_RATIO)

```

We want to see for each condition x bin x gene how many simulations there are:

```{r}
#range is from 165 to 10,000 simulations
num_sims = filtered_sims_wider %>%
  select(STRAIN, TISSUE, AGE_BIN, BIN, GENE) %>%
  group_by(STRAIN, TISSUE, AGE_BIN, BIN, GENE) %>%
  summarise(SIM_COUNT = n())

```

We export the sim_hnhs ratios

```{r}
setwd("/Users/isabelserrano/Documents/Science/Analyses/Conplastic_Strains/files_and_analyses/genomewide_selection_scan/hnhs_per_freq_bin/")

write.table(sim_ratios, file = paste(outdir_files,"/sim_hNhS_ratios_per_gene_per_bin.txt", sep = ""), sep = "\t", quote = F, row.names = T)

```

